\textbf{COMP0085 Summative Assignment}

### 2022-12-30 submission.tex
update mean field

@@

@@

@@

@@

%\[\mathcal{M}_{j \rightarrow i}(s_i) = \tilde{g}_{ji, \neg s_j}(s_i)\]

%\[\mathcal{M}_{j \rightarrow i} = \tilde{f}_j(s_j) \prod_{k\in ne(j), k\neq i}^{K}}\tilde{g}_{kj, \neg s_k}(s_j)  \sum_{\substack{s_m \in \{0, 1\} \\ m \in \{1, ..., j-1, j+1,... K\}}} \prod_{k\in ne(j), k\neq i}^{K}} \tilde{g}_{kj, \neg s_j}(s_k)\]

%\[\mathcal{M}_{j \rightarrow i}(s_j) = \tilde{f}_j(s_j) \prod_{k\in ne(j), k\neq i}^{K}}\tilde{g}_{kj, \neg s_k}(s_j) \prod_{k\in ne(j), k\neq i}^{K}} \left(\sum_{s_k \in \{0, 1\} } \tilde{g}_{kj, \neg s_j}(s_k) \right)\]
%
%\[\mathcal{M}_{j \rightarrow i}(s_j) = \tilde{f}_j(s_j) \prod_{k\in ne(j), k\neq i}^{K}}\tilde{g}_{kj, \neg s_k}(s_j)\]

@@

%\[
%\begin{array}{l}
% \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \log \left(\frac{\theta_i}{1-\theta_i} \right) s_i + \log \left(\frac{\theta_j}{1-\theta_j} \right) s_j\\
%\tab \tab \tab \tab \tab + \sum_{i'=1}^{K}  \log \left(\frac{\lambda_{i'}}{1-\lambda_{i'}} \right) s_{i'} \\
%\tab \tab \tab \tab \tab +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_{i'}}{1-\theta_{i'}} \right) s_{i'} + \log \left(\frac{\theta_{j'}}{1-\theta_{j'}} \right) s_{j'}
%\end{array}
%\]

%Only including terms with $s_i$ and $s_j$:
%
%\[
%\begin{array}{l}
% \log \tilde{g}_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \left((K-1) \log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i\\
%\tab \tab \tab \tab \tab +\left(
%(K-1) \log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
%\end{array}
%\]

@@

%\[
%\begin{array}{l}
% \log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\\
%\tab \tab \tab \tab \tab + \sum_{i'=1}^{K}  \log \left(\frac{\lambda_{i'}}{1-\lambda_{i'}} \right) s_{i'} \\
%\tab \tab \tab \tab \tab +  \sum_{i'=1}^{K}\sum_{\substack{j'=1 \\ i', j' \neq i, j}}^{i'-1}  \log \left(\frac{\theta_{i'}}{1-\theta_{i'}} \right) s_{i'} + \log \left(\frac{\theta_{j'}}{1-\theta_{j'}} \right) s_{j'}
%\end{array}
%\]

%Only including terms with $s_i$ and $s_j$:
%
%\[
%\begin{array}{l}
% \log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto \frac{- \mu_i^T \mu_j}{\sigma^2} s_i s_j\\
%\tab \tab \tab \tab \tab + \left((K-2)\log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\right) s_i \\
%\tab \tab \tab \tab \tab +  \left((K-2)\log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\right) s_j
%\end{array}
%\]
%
%We can write this as:
%
%\[
%\log g_{ij}(s_i, s_j) q_{\neg \tilde{g}_{ij}(s_i, s_j)}(\textbf{s}) \propto W_{ij} s_i s_j+ A_{s_i} s_i + B_{s_j} s_j
%\]
%
%where:
%
%\[A_{s_i} = (K-2)\log \left(\frac{\theta_i}{1-\theta_i} \right) + \log \left(\frac{\lambda_i}{1-\lambda_i} \right)\]
%
%and
%
%\[B_{s_j} = (K-2)\log \left(\frac{\theta_j}{1-\theta_j} \right) + \log \left(\frac{\lambda_j}{1-\lambda_j} \right)\]

@@

Our message passing approximations:

\[\exp(\eta_{ij} s_i) = \tilde{f}_{i}(s_{i}) \prod_{k\in ne(i), k\neq j}^{K} \mathcal{M}_{k \rightarrow i} \]

\[\exp(\eta_{ji} s_j) = \tilde{f}_{j}(s_{j}) \prod_{k\in ne(j), k\neq i}^{K} \mathcal{M}_{k \rightarrow j}\]

where each message $\mathcal{M}_{j \rightarrow i}$ has a factored approximation:

\[\mathcal{M}_{k \rightarrow i} = \exp(\eta_{ki} s_k) \]

because each site $\tilde{g}_{jk}(s_j s_k)$ is approximated as a product of two messages $\mathcal{M}_{j \rightarrow k}\mathcal{M}_{k \rightarrow j}$, each a Bernoulli.

Thus, the natural parameters of the messages are updated with:

\[\eta_{ij} = b_i + \sum_{k\in ne(i), k\neq j}^{K} \eta_{ki} \]

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{outputs/q6/all-latent-factors}
\caption{Latent factors learned with EP/Loopy-BP}
\label{fig:6-latent-factors}
\end{figure}

>>>>

\newpage
\subsection*{(e)}

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{outputs/q3/f-free-energy}
\caption{Free Energy}
\label{fig:3f-free-energy}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{outputs/q3/f-init-latent-factors}
\caption{Initial Latent Factors}
\label{fig:3f-init-latent-factors}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{outputs/q3/f-latent-factors}
\caption{Learned Latent Factors}
\label{fig:3f-latent-factors}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{outputs/q3/g-free-energy-diff-sigma}
\caption{Free energy convergence vs $\sigma$}
\label{fig:3g-free-energy-diff-sigma}
\end{figure}

@@

@@

@@

@@

Using factored approximate messages, we see that:

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{outputs/q6/all-init-latent-factors}
\caption{Initial Latent factors learned with EP/Loopy-BP}
\label{fig:6-init-latent-factors}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[scale=0.4]{outputs/q6/all-latent-factors}
\caption{Learned Latent factors learned with EP/Loopy-BP}
\label{fig:6-latent-factors}
\end{figure}
